A Computational Account of Default Mode Function\\ by Reinforcement Learning Theory

### 2017-03-26 paper/paper.tex
interim

The ventromedial prefrontal cortex (vmPFC) subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adapted behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on model-based value representations
\citep{doherty2015structure}.
Quantative lesion findings across 344 human individuals confirmed
a gross impairment in value-based decision making
\citep{glascher2012lesion}.
The vmPFC is preferentially connected with limbic and reward-related areas.
The vmPFC has been observed to have monosynaptical connections
with the nucleus accumbens
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
correlated with indices of value-guided behavior and social reward attitudes
\citep{lebreton2009automatic}.
The NAc is thought to be linked to reward mechanisms that not
only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly.
This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified this brain regions to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the nucleus accumbens, hippocampus (HC),
and posteriomedial cortex.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in OFC.
Such a \textit{cognitive map} of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume GMV of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity,
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.
%

\subsection{The hippocampus: Memory, space, and experience replay}
The HC is well known to be involved in memory and spatial navigation in animals and humans.
Its highly recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
%
While the HC in the medial temporal lobe system
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering}.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one’s own future and
imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
posterior hippocampus \citep{maguire2000navigation}.
Indeed, place-specific neuronal populations in the hippocampus
spike one after another when an animal is choosing between alternative
paths \citep{johnson2007neural}; and
these neuronal patterns apear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

@@

\subsection{Value functions and policies}
The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ?}
  The amygdala is known to be involved in significance evaluation) and
  the nucleus accumbens (NAc) is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
as implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus the agent is positively biased towards considering rewards which are will
come soon. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

@@

\begin{itemize}
  \item The policy matrix represents the repertoire of possible actions
  on the world given a current state. It encodes the probabilites of
  choosing actions to be executed in a certain situation.
  \item Interactions between external and internal sources of
  information are probably commonplace during perception
  \item The MDP frameworks how the human brain could solve
  environmental problems over extended time periods.
  \item Given a state, the value function represents the estimated utility of
  being in that state, which is subjective to the agent.
  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
  "How desirable is it to be in state $s$ ?"
  For each action $a$ that can be taken from state $s$,
  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
  utility of executing a specific action; it answers the question
  "What is the expected utiltiy of taking action $a$ in this situation ?"
  \item The DMN may subserve
  constant exploration of possible future actions and their
  cumulative outcomes. Implicit computation of future choices
  provides an explanation for purposeful mind-wandering.
  \item hippocampus -> provides perturbed action-transition-state-reward samples
  as a block of "imagined", "hypothesized", "recalled" experience;
  the small variations in these experience blocks allow searching
  a larger space of parameters and possible experiences.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing possible
  future scenarios and action outcomes.
  Our approach acknowledges the unavoidable stochasticity of
  computation in neuronal processes \citep{faisal2008noise}.
  \item
  \textit{Inference} in the human brain reduces to generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  \item In the RL view, plasticity in the DMN arises naturally.
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy matrix
  and value function until the system converges to a
  new steady-state of optimal action decisions given certain states.
  \item The DMN is today known to consistently increase in neural
  activity when engaging cognitive processes that are detached from
  current sensory environment. DMN involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the
  MDP's implicit computation of action and state cascades as a function
  of what happened in the past.
\end{itemize}

@@

\section{Relation to other statistical models (Elvis)}
\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
\textbf{XXX: add tons of more examples and connnections}
\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
In summary:
\begin{itemize}
  \item Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute with the free energy principle
  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
  backprop!
  \end{itemize}

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding framework
\citep{clark2013whatever}
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

@@

\subsection{Semantic Hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extractable from the environment
\citep{bar2009proactive}.
%
Going much beyond human language ifself,
semantic building blocks provide the basis to
mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

@@

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\citep{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\citep{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and the complex relationships between them.

@@

\section{Conclusion}
What brain function could be important enough
for the existence and survival of the human species
to warrant constant, high energy costs?
MDPs provide an attractive
computational model how the human association cortex
might implement supra-modal representation and control of the environment to
optimize the organism's fate.
% by obtaining reward and avoiding pain.
This idealized process model explains
a number of previous experimental observations in the
DMN by a simple but non-trivial mechanism.
%
Viewed as an MDP, behavior unfolds by integrating happened past events
and possible future events to guide action choice in the present context.
This functional account is more compatible with the DMN's
poorly understood involvement across
autobiographical memory retrieval, problem solving,
abstract reasoning based on internally generated scenes, social cognition,
as well as delay discounting and self-related prospection.
MDPs provide a mathematical formalism how
optimal substructure in the environment can be recursively exploited
when confronted with complicated decisions.
Algorithmic gains by injecting stochasticity in the recall of past
actions and outcomes to improve the internal world model may explain why
very accurate memories have been disfavored in human evolution
and why human creativity might be adaptive.
%
In principle,
neuroscientific experiments can be designed that operationalize
the set of action, value, and state variables that determine
the behavior of intelligent RL systems.
The proposed machine-learning perspective
on DMN function is hence not only practically computable but
also neuroscientifically falsifiable.
At the least, we propose an alternative vocabulary to
describe and interpret neuroscientific studies in DMN research.
%
Ultimately,
the DMN can be viewed as a smoothing kernel extending
from the relatively recent events to anticipated upcoming events
to constantly improve sensory processing and
action performance in the world.

>>>>

The ventromedial prefrontal cortex (vmPFC) subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adapted behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on model-based value representations
\citep{doherty2015structure}.
Quantative lesion findings across 344 human individuals confirmed
a gross impairment in value-based decision making
\citep{glascher2012lesion}.
The vmPFC is preferentially connected with limbic and reward-related areas.
The vmPFC has been observed to have monosynaptical connections
with the nucleus accumbens
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
correlated with indices of value-guided behavior and social reward attitudes
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified this brain regions to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the nucleus accumbens, hippocampus (HC),
and posteriomedial cortex.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in OFC.
Such a \textit{cognitive map} of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume GMV of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity,
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.
%

\subsection{The hippocampus: Memory, space, and experience replay}
The HC is well known to be involved in memory and spatial navigation in animals and humans.
Its highly recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
%
While the HC in the medial temporal lobe system
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally}.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one’s own future and
imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
posterior hippocampus \citep{maguire2000navigation}.
Indeed, place-specific neuronal populations in the hippocampus
spike one after another when an animal is choosing between alternative
paths \citep{johnson2007neural}; and
these neuronal patterns apear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

@@

\subsection{Value functions and policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ?}
  The amygdala is known to be involved in significance evaluation) and
  the nucleus accumbens (NAc) is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
as implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus the agent is positively biased towards considering rewards which are will
come soon. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

@@

\begin{itemize}
  \item The policy matrix represents the repertoire of possible actions
  on the world given a current state. It encodes the probabilites of
  choosing actions to be executed in a certain situation.
  \item Interactions between external and internal sources of
  information are probably commonplace during perception \citep{gershman2010learning}
  \item The MDP frameworks how the human brain could solve
  environmental problems over extended time periods.
  \item Given a state, the value function represents the estimated utility of
  being in that state, which is subjective to the agent.
  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
  "How desirable is it to be in state $s$ ?"
  For each action $a$ that can be taken from state $s$,
  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
  utility of executing a specific action; it answers the question
  "What is the expected utiltiy of taking action $a$ in this situation ?"
  \item The DMN may subserve
  constant exploration of possible future actions and their
  cumulative outcomes. Implicit computation of future choices
  provides an explanation for purposeful mind-wandering.
  \item hippocampus -> provides perturbed action-transition-state-reward samples
  as a block of "imagined", "hypothesized", "recalled" experience;
  the small variations in these experience blocks allow searching
  a larger space of parameters and possible experiences.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing possible
  future scenarios and action outcomes.
  Our approach acknowledges the unavoidable stochasticity of
  computation in neuronal processes \citep{faisal2008noise}.
  \item
  \textit{Inference} in the human brain reduces to generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  \item In the RL view, plasticity in the DMN arises naturally.
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy matrix
  and value function until the system converges to a
  new steady-state of optimal action decisions given certain states.
  \item The DMN is today known to consistently increase in neural
  activity when engaging cognitive processes that are detached from
  current sensory environment. DMN involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the
  MDP's implicit computation of action and state cascades as a function
  of what happened in the past.
\end{itemize}

@@

\section{Relation to other statistical models (Elvis)}
\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
\textbf{XXX: add tons of more examples and connnections}
\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
In summary:
\begin{itemize}
  \item Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute with the free energy principle
  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
  backprop!
  \end{itemize}

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding framework
\citep{clark2013whatever, friston2008hierarchical}
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

@@

\subsection{Semantic Hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extractable from the environment
\citep{bar2009proactive}.
In this way, the encoding of complex environments could be facilitated
by relative similarity associations between states.
%
Going much beyond human language ifself,
semantic building blocks provide the basis to
mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

@@

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\citep{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\citep{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and
the complex relationships between them
\citep{schilbach2008minds}.

@@

\section{Conclusion}
What brain function could be important enough
for the existence and survival of the human species
to warrant constant, high energy costs?
MDPs provide an attractive
computational model how the human association cortex
might implement supra-modal representation and control of the environment to
optimize the organism's fate.
% by obtaining reward and avoiding pain.
This idealized process model explains
a number of previous experimental observations in the
DMN by a simple but non-trivial mechanism.
%
From the computational viewed of a MDP,
behavior unfolds by integrating happened past events
and possible future events to guide action choice in the present context.
This functional account is more compatible with the DMN's
poorly understood involvement across
autobiographical memory retrieval, problem solving,
abstract reasoning based on internally generated scenes, social cognition,
as well as delay discounting and self-related prospection.
MDPs provide a mathematical formalism how
optimal substructure in the environment can be recursively exploited
when confronted with complicated decisions.
Algorithmic gains by injecting stochasticity in the recall of past
actions and outcomes to improve the internal world model may explain why
very accurate memories have been disfavored in human evolution
and why human creativity might be adaptive.
%
In principle,
neuroscientific experiments can be designed that operationalize
the set of action, value, and state variables that determine
the behavior of intelligent RL systems.
The proposed machine-learning perspective
on DMN function is hence not only practically computable but
also neuroscientifically falsifiable.
At the least, we propose an alternative vocabulary to
describe and interpret neuroscientific studies in DMN research.
%
Ultimately,
the DMN can be viewed as a smoothing kernel extending
from the relatively recent events to anticipated upcoming events
to constantly improve sensory processing and
action performance in the world.
